TINET+TECS: Component-Based TCP/IP Protocol Stack for Embedded Systems

### 2017-06-22 ICESS2017/paper.tex
Final version
Camera-Ready ver.

This paper evaluates the overheads of execution time and memory consumption and the amount of code line changes needed to add and remove functionalities to improve configurability with small overheads.
Furthermore, the advantages of dynamic connection in terms of the memory consumption and low overhead of the TECS adapter are demonstrated.

@@

TINET is a compact TCP/IP protocol stack for embedded systems based on the ITRON\footnote{ITRON is a realtime operating system (RTOS) developed by the TRON project.} TCP/IP API Specification \cite{url:ITRON_TCP/IP_API_Spec}, developed by the TOPPERS Project \cite{url:TOPPERS}.
TINET has been released as an open-source tool.

@@

TINET+TECS utilizes the dynamic connection to switch between CEP and REP components, as shown in Fig. \ref{fig:DynamicConnectionUseCase}.
In a server application, CEP is associated with REP in the state of waiting for connection request from clients\footnote{tcp\_acp\_cep(ID cepid, ID repid, T\_IPV4EP *p\_dstaddr, TMO tmout).}.
For example, when processing with the HTTP protocol, CEP passively opens with an REP of port number 80.

@@

\begin{figure}[t]
\centering
\begin{lstlisting}
signature sRepSelector {
    void  getRep([out]Descriptor(sREP4) *desc,
                 [in]int_t i);
};
celltype tRepSelector {
    entry sRepSelector eRepSelector;
    [ref_desc]
        call sREP4 cREP[NUM_REP];
};
celltype tTCPCEP {
    call sRepSelector cRepSelector;
    [dynamic]
        call sREP4 cREP;
    /* Omit: other call/entry ports */
    /* Omit: attributes and variables */
};
\end{lstlisting}
\vspace{-2mm} \caption{Signature and celltype description for dynamic connection}
\vspace{-1mm} \label{src:DynamicCDLcode}
\end{figure}

Fig. \ref{src:DynamicCcode} shows a sample code of dynamic connection.
The eAPI\_accept function is the function wrapping {\it tcp\_acp\_cep} under TECS, which is set as the state waiting for connection request.
Dynamic connection in the function is performed as shown in Fig. \ref{src:DynamicCcode}.
First, the descriptor of REP to be joined is obtained (Line 3 in Fig. \ref{src:DynamicCcode}).
The first argument, {\it \&desc}, is a variable used to store the descriptor information, whereas the second argument, {\it repid}, is the index of tREP cells.
Next, the descriptor is set (Line 5 in Fig. \ref{src:DynamicCcode}), and the cREP port combines the tREP cell specified by the descriptor, enabling the tCEP cell to call the function of the tREP cell to be joined (Line 7 in Fig. \ref{src:DynamicCcode}).

\begin{figure}[t]
\centering
\begin{lstlisting}
eAPI_accept (.., ..) {
    /* Get a descriptor of intended REP cell */
    cRepSelector_getRep(&desc, repid);
    /* Set the descriptor */
    cREP_set_descriptor(desc);
    /* Call the function of intended REP cell */
    cREP_getEndpoint();
}
\end{lstlisting}
\vspace{-2mm} \caption{Accept function (Dynamic connection example)}
\vspace{-1mm} \label{src:DynamicCcode}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8.0cm,clip]{figure/ComparisonOfDynamicConnection.pdf}
    \vspace{-1mm} \caption{Component diagrams for without/with dynamic connection cases}
    \vspace{-1mm} \label{fig:ComparisonOfDynamicConnection}
\end{figure}

Memory consumption without and with TECS dynamic connection was then evaluated.
As shown in the left panel of Fig. \ref{fig:ComparisonOfDynamicConnection}, each CEP component should be statically connected to all REP components if dynamic connection is not used.
As the number of REPs increases, additional call ports of CEP are required, in turn increasing the consumption of memory. 
Dynamic connection reduces memory consumption because only one CEP-to-REP call port is required per CEP, as illustrated with red lines in the right panel of Fig. \ref{fig:ComparisonOfDynamicConnection}.
Even if the number of REPs increases, additional call ports can be joined through the selector, instead of the CEPs.

@@

The code lines in CDL of without and with dynamic connection is shown in TABLE \ref{tab:EvaluationOfConfigurabilityByDynamicConnection} to demonstrate improved configurability.
As the number of CEPs and REPs increases, the amount of CDL code lines to be added increases.
In the left panel of Fig. \ref{fig:ComparisonOfDynamicConnection}, each CEP connects all REPs as shown in the upper of Fig. \ref{src:ComparisonOfCDL}. 
In the right panel of Fig. \ref{fig:ComparisonOfDynamicConnection}, a CEP dynamically connects an REP, and only the selector connects all REPs as shown in the lower of Fig. \ref{src:ComparisonOfCDL}. 
It is effective for software that uses many ports because the difference spreads as the number of CEPs and REPs increases.

@@

\begin{table*}[t]
    \caption{Comparison of related and proposed framework}
    \label{tab:comparison}
\vspace{-2mm}
    \centering
    \begin{tabular}{c||ccccccc}
        \hline\hline
        & \shortstack{Configurability}
        & \shortstack{Generation} 
        & \shortstack{Binding} 
        & \shortstack{Support CBD} 
        & \shortstack{Memory saving} \\ \hline
        uIP \cite{par:uIP}     & $\bigtriangleup$          & Dynamic    & Dynamic  &              & \checkmark \\
        lwIP \cite{par:lwIP}   & $\bigtriangleup$          & Dynamic    & Dynamic  &              & \checkmark \\
        TINET \cite{url:TINET} &            & Static     & Dynamic  &              & \checkmark \\
        Proposed framework     & \checkmark & Static     & Dynamic  & \checkmark   & \checkmark \\
    \end{tabular}
        \vspace{-5mm}
\end{table*}

@@

This paper also presented dynamic connection, a new TECS method, to enable dynamic processing while reducing memory consumption.
TINET+TECS utilizes dynamic connection to satisfy the TINET specification for supporting static generation of CEPs and REPs.
As the TECS adapter supports legacy codes, existing TCP/IP applications can run without modification in the proposed framework.
TINET+TECS and application programs used in the evaluation are all open-source, and may be downloaded from our website \cite{url:TINET+TECS}.

@@

The authers would like to thank Hiroaki Nagashima for supporting this research.
This work is supported by JSPS KAKENHI Grant Number 15H05305.

>>>>

This paper evaluates the overheads of execution time and memory consumption and the amount of code line changes needed to add and remove functionalities to improve configurability with small overheads.
Furthermore, the advantages of the dynamic connection in terms of the memory consumption and low overhead of the TECS adapter are demonstrated.

@@

TINET is a compact TCP/IP protocol stack for embedded systems based on the ITRON\footnote{ITRON is a real-time operating system (RTOS) developed by the TRON project.} TCP/IP API Specification \cite{url:ITRON_TCP/IP_API_Spec}, developed by the TOPPERS Project \cite{url:TOPPERS}.
TINET has been released as an open-source tool.

@@

TINET+TECS utilizes the dynamic connection to switch between CEP and REP components, as shown in Fig. \ref{fig:DynamicConnectionUseCase}.
In a server application, CEP is associated with REP in the state of waiting for a connection request from clients\footnote{tcp\_acp\_cep(ID cepid, ID repid, T\_IPV4EP *p\_dstaddr, TMO tmout).}.
For example, when processing with the HTTP protocol, CEP passively opens with an REP of port number 80.

@@

\begin{figure}[t]
\centering
\begin{lstlisting}
signature sRepSelector {
    void  getRep([out]Descriptor(sREP4) *desc,
                 [in]int_t i);
};
celltype tRepSelector {
    entry sRepSelector eRepSelector;
    [ref_desc]
        call sREP4 cREP[NUM_REP];
};
celltype tTCPCEP {
    call sRepSelector cRepSelector;
    [dynamic]
        call sREP4 cREP;
    /* Omit: other call/entry ports */
    /* Omit: attributes and variables */
};
\end{lstlisting}
\vspace{-2mm} \caption{Signature and celltype description for the dynamic connection}
\vspace{-1mm} \label{src:DynamicCDLcode}
\end{figure}

Fig. \ref{src:DynamicCcode} shows a sample code of the dynamic connection.
The eAPI\_accept function is the function wrapping {\it tcp\_acp\_cep} under TECS, which is set as the state waiting for a connection request.
The dynamic connection in the function is performed as shown in Fig. \ref{src:DynamicCcode}.
First, the descriptor of REP to be joined is obtained (Line 3 in Fig. \ref{src:DynamicCcode}).
The first argument, {\it \&desc}, is a variable used to store the descriptor information, whereas the second argument, {\it repid}, is the index of tREP cells.
Next, the descriptor is set (Line 5 in Fig. \ref{src:DynamicCcode}), and the cREP port combines the tREP cell specified by the descriptor, enabling the tCEP cell to call the function of the tREP cell to be joined (Line 7 in Fig. \ref{src:DynamicCcode}).

\begin{figure}[t]
\centering
\begin{lstlisting}
eAPI_accept (.., ..) {
    /* Get a descriptor of intended REP cell */
    cRepSelector_getRep(&desc, repid);
    /* Set the descriptor */
    cREP_set_descriptor(desc);
    /* Call the function of intended REP cell */
    cREP_getEndpoint();
}
\end{lstlisting}
\vspace{-2mm} \caption{Accept function (a dynamic connection example)}
\vspace{-1mm} \label{src:DynamicCcode}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8.0cm,clip]{figure/ComparisonOfDynamicConnection.pdf}
    \vspace{-1mm} \caption{Component diagrams for without/with the dynamic connection cases}
    \vspace{-1mm} \label{fig:ComparisonOfDynamicConnection}
\end{figure}

Memory consumption without and with TECS dynamic connection was then evaluated.
As shown in the left panel of Fig. \ref{fig:ComparisonOfDynamicConnection}, each CEP component should be statically connected to all REP components if the dynamic connection is not used.
As the number of REPs increases, additional call ports of CEP are required, in turn increasing the consumption of memory. 
The dynamic connection reduces memory consumption because only one CEP-to-REP call port is required per CEP, as illustrated with red lines in the right panel of Fig. \ref{fig:ComparisonOfDynamicConnection}.
Even if the number of REPs increases, additional call ports can be joined through the selector, instead of the CEPs.

@@

The code lines in CDL of without and with the dynamic connection is shown in TABLE \ref{tab:EvaluationOfConfigurabilityByDynamicConnection} to demonstrate improved configurability.
As the number of CEPs and REPs increases, the amount of CDL code lines to be added increases.
In the left panel of Fig. \ref{fig:ComparisonOfDynamicConnection}, each CEP connects all REPs as shown in the upper of Fig. \ref{src:ComparisonOfCDL}. 
In the right panel of Fig. \ref{fig:ComparisonOfDynamicConnection}, a CEP dynamically connects an REP, and only the selector connects all REPs as shown in the lower of Fig. \ref{src:ComparisonOfCDL}. 
It is effective for software that uses many ports because the difference spreads as the number of CEPs and REPs increases.

@@

\begin{table*}[t]
    \caption{Comparison of related and proposed framework}
    \label{tab:comparison}
\vspace{-2mm}
    \centering
    \begin{tabular}{c||ccccccc}
        \hline\hline
        & \shortstack{Configurability}
        & \shortstack{Generation} 
        & \shortstack{Binding} 
        & \shortstack{Support CBD} 
        & \shortstack{Memory saving} \\ \hline
        uIP \cite{par:uIP}     & $\bigtriangleup$          & Dynamic    & Dynamic  &              & \checkmark \\
        lwIP \cite{par:lwIP}   & $\bigtriangleup$          & Dynamic    & Dynamic  &              & \checkmark \\
        TINET \cite{url:TINET} &            & Static     & Dynamic  &              & \checkmark \\
        Proposed framework     & \checkmark & Static     & Dynamic  & \checkmark   & \checkmark \\
    \end{tabular}
\end{table*}

@@

This paper also presented the dynamic connection, a new TECS method, to enable dynamic processing while reducing memory consumption.
TINET+TECS utilizes the dynamic connection to satisfy the TINET specification for supporting the static generation of CEPs and REPs.
As the TECS adapter supports legacy codes, existing TCP/IP applications can run without modification in the proposed framework.
TINET+TECS and application programs used in the evaluation are all open-source, and may be downloaded from our website \cite{url:TINET+TECS}.

@@

The authors would like to thank Hiroaki Nagashima for supporting this research.
This work is supported by JSPS KAKENHI Grant Number 15H05305.
